---
title: "医学用生存分析(单因素)"
subtitle: "——肿瘤数据例子"
author: "3gee"
date: "2022-10-17"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{css, echo=FALSE}
p, h1, h2, h3, h4, h5 {
  font-family:"Microsoft YaHei",微软雅黑;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_ALL", "CHS")
library(survival)
library(survminer)
library(tidyverse)
```

## 简介 | Introduction

本`RMarkdown`用来展示如何用`R`实现**生存分析**。

This `RMarkdown` file shows how to use `R` to analysis survival data.


## 读入数据和整理 | Data Loading and Transfering

```{r}
df <- xlsx::read.xlsx("../datasets/肿瘤大小生存数据.xlsx", 1)

test_df <- df %>%
  mutate(is_deleted = `删失`) %>%
  mutate(group = ifelse(`肿瘤大小` == '小于3cm', 0, 1)) %>%
  mutate(month = `生存月数`)

test_df
```

## 建立模型 | Design a Survival Model

```{r}
m <- survfit(Surv(month, is_deleted) ~ group, data=test_df)
m
``` 
```{r}
plot(m)
```
```{r}
ggsurvplot(m,surv.median.line = "hv",)
```

The survival chance is 1.0 at time zero (or 100 percent of the participants are alive).

At time 20, the chances of survival for group=1 are about 0.50 (or 50 percent) and 0.85 (or 85 percent) for group=0.

The median survival time for group=1 is 20 days and for group=0 is 40 days, indicating that those who has small tumor size(less then 3cm) has a better survival rate than those has large tumor size.


```{r}
summary(m)$table
```

### LogRank测试 | LogRank Testing

LogRank Testing is used to check the significance of difference between two survival model.


```{r}
survdiff(Surv(month, is_deleted) ~ group, data=test_df)
```

The Chi-Squared test statistic is 7.4 with 1 degree of freedom and the corresponding p-value is 0.007. Since this p-value is less than 0.05, we reject the null hypothesis.

In other words, we have sufficient evidence to say that there is a statistically significant difference in survival between the two groups.
